### 2009-09-20 latex/paper1/contents.tex
paper

@@

% Neeraj Sangal
% Dependencies are extracted from the code by a conventional static analysis
%
% Ao falar de dependências, cita
%   Designing Software for Ease of Extension and Contraction
%
%The LDM tool uses a standard notion of dependency, in which a
%module A depends on a module B if there are explicit references
%in A to syntactic elements of B. Currently, a module is a Java
%class, but a more fine-grained analysis is possible. This simple but
%effective notion of dependency works well for understanding
%design dependencies, in which modifications to one module might XXX
%affect another. It is less well suited to determining runtime
%properties (such as how failures can propagate between modules),
%which require a deeper static analysis.
%
%---------
%
% Parnas
%
% A uses B if correct execution of B may be necessary for A to complete the
% task described in its specification.
%
% A uses B if there exist situations in which the correct functioning of A
% depends upon the availability of a correct implementation of B.
%
% "Uses" can also be formulated as "requires the presence of a correct version
% of".
%
% Uses vs. invokes

% TODO: CITACAO DEPOIS DO CONCEITO
In the context of our research, we consider classes as the components

@@

One of the most obvious ways to reduce the time needed to develop or maintain a
software system is to assign more developers to it. When combined with poor
management, though, adding developers may in fact raise the development time,
due to the increasing communication costs between developers \cite{Brooks1995}.

The communication costs are affected by the assignment of developers to
implementation components of a system.
%The assignment of developers to implementation components of a system affects
%the amount of communication needed between developers. 
If two development teams are responsible for maintaining components that are
highly interdependent, they need to coordinate their work often in order avoid
code duplication and breaking each other's component funcionality.

Therefore, to optimize the development, the manager can decompose the system
into modules, or groups of components, so that there are as few inter-module
dependencies as possible (i.e., the modules are weakly coupled). Each module is,
then, threated as an unit of work assignment \cite{Parnas1972}: each module is
assigned to one team and, because of the little inter-module dependency, the
communication between teams is reduced, therefore speeding up the development.

To decompose a system into weakly coupled modules, it is necessary to have a
global knowledge of the dependencies between its components. Although dependency
analysis tools can provide such information, the task of finding modules can be
overwhelming because of the large number of components and dependencies.

The task can be made easier by the use of software clustering algorithms, also
known as also known as architecture recovery algorithms \cite{Pollet2007}. They
automatically decompose a software system into weakly coupled modules by
analyzing the network of dependencies between its components.

%This kind of network is the input for many software clustering
%algorithms \cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005}.

Many algorithms were proposed, but there is little information concerning their
empirical evaluation.

% sw clustering is a task of reverse engineering \cite{Tonella2007}.

An important question for the adoption of a software clustering algorithm is
whether the decomposition it finds for a system is similar to a reference
decomposition for that system, i.e., a decomposition found by a group of
experienced developers. To answer this question one needs to test the algorithm
by applying it to a sample of systems with known reference decompositions, and
then compare the two decompositions for each system by means of a metric such as
MoJo \cite{Tzerpos1999} and EdgeSim \cite{Mitchell2001}.

Unfortunately there are few systems with known reference decompositions.
Furthermore, because such decompositions are costly do obtain
\cite{Koschke2000}, there are few empirical studies about software clustering
algorithms, most of them based on a couple of small and medium systems
\cite{Anquetil1999,Maqbool2007,Bittencourt2009}.

@@

In this research we propose to evaluate software clustering algorithms by
applying them to synthetic, i.e., computer generated, component dependency
networks with built-in decompositions. With this approach it is feasible to test
algorithms with arbitrarily large samples in a controlled manner.

Of course, it is desirable that the synthetic networks resemble networks
extracted from real software systems. Hence, in this paper, we describe a study
about three models that synthesize networks with built-in decompositions. We
show empirically that all three models are capable of synthesizing networks that
resemble software networks.

% TODO: antropomorphic sections!
The remaining sections are organized as follows:
Section \ref{sec:swnets} is a overview of the main concepts used in this paper.
%
Section \ref{sec:models} presents three network models.
%
Section \ref{sec:characterization} demonstrates a method to tell if a network
resembles a software network.
%
Section \ref{sec:evaluation} describes an experiment that shows that the three
models synthesize networks that resemble software networks.
%
Section \ref{sec:conclusion} discusses the impact of this research and future
work.

%alike, resembling, exchangeable, indistiguible

%\section{Related Work}
%
%\section{Software Modularization Recovery: an Overview}
%Aka software architecture recovery, software architecture reconstruction,
%software clustering...
%It's an task of reverse engineering, as defined by Tonella \cite{Tonella2007}...
%DEFINITIONS
%external edge...

%\section{Software Networks and Software Clustering}
%
%directed graph, (un)weighted

\section{Software Networks} \label{sec:swnets}
% TODO: consider merging with introduction

At the implementation level, a software system can be viewed as a network of
interacting components. In a object-oriented system, for example, classes
interact with other classes through mechanisms such as inheritance and
aggregation. By abstracting the particular mechanisms of interaction, it is
possible to build a network of dependencies between components, represented as a
directed graph. This kind of network is the input for many software clustering
algorithms \cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005}.

In this paper we will study networks of dependencies between classes in
object-oriented systems. We will call any network of this kind a \emph{software
network}. This representation of a software system is amenable to analysis
through network theory. %Software networks can be studied by network theory.

%Define: Component. Component dependency network. Directed unweighted graph.
%Clustering based on dependencies.

%(in fact we can't argue the degree distribution is perfectly fit by a power law.
%Anyway, the distribution is much more assymetric than normal ou Poisson's)
%
%software indegree distribution (power law), outdegree distribution (not so power
%law)

@@

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, it has been found that many networks
arising from sociology, biology, technology and other domains present remarkable
structural similarities. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{\-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to software networks and it was shown that they
are also scale-free networks \cite{Myers2003,Valverde2002}. Thus, the study of
the structure of software systems in general can be benefited from the vast
literature available about scale-free networks.

Many models were proposed to explain the formation of scale-free networks. These
models are algorithms that can be proven, either formally or empirically, to
synthesize networks that are scale-free. Most models, though, do not account for
modular structure. In this section we present three models that synthesize
directed networks with built-in modular decompositions: BCR+, CGW and LF.

%The first two models grow networks using a mechanism called
%preferential attachment CITE Barabasi, by which nodes with many edges tend to
%receive more edges in the process.

%\subsection{ER (Erdos-Renyi)}

%\subsection{Configuration model}
% take the most typical software system and steal its degree sequence

@@

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be too low, because most software networks would be misclassified.
So we chose the value 0.88, that is immediatly greater than the second greater
non-software network S-score. With this value, we have both high recall (95.4\%)
and high precision (96.9\%).

>>>>

% TODO: say that we use DIRECTED GRAPHS

@@

% Neeraj Sangal
% Dependencies are extracted from the code by a conventional static analysis
%
% Ao falar de dependências, cita
%   Designing Software for Ease of Extension and Contraction
%
%The LDM tool uses a standard notion of dependency, in which a
%module A depends on a module B if there are explicit references
%in A to syntactic elements of B. Currently, a module is a Java
%class, but a more fine-grained analysis is possible. This simple but
%effective notion of dependency works well for understanding
%design dependencies, in which modifications to one module might
%affect another. It is less well suited to determining runtime
%properties (such as how failures can propagate between modules),
%which require a deeper static analysis.
%
%---------
%
% Parnas
%
% A uses B if correct execution of B may be necessary for A to complete the
% task described in its specification.
%
% A uses B if there exist situations in which the correct functioning of A
% depends upon the availability of a correct implementation of B.
%
% "Uses" can also be formulated as "requires the presence of a correct version
% of".
%
% Uses vs. invokes

@@

% TODO: this paragraph can mostly disappear.
%
The time spent on the maintenance of a software system can be affected by the
assignment of developers to implementation components of the system.
If two development teams are responsible for maintaining components that are
highly interdependent, they need to coordinate their work often in order avoid
code duplication and breaking each other's component funcionality.

%One of the most obvious ways to reduce the time needed to develop or maintain a
%software system is to assign more developers to it. When combined with poor
%management, though, adding developers may in fact raise the development time,
%due to the increasing communication costs between developers \cite{Brooks1995}.
%
%The communication costs are affected by the assignment of developers to
%implementation components of a system.
%%The assignment of developers to implementation components of a system affects
%%the amount of communication needed between developers. 
%If two development teams are responsible for maintaining components that are
%highly interdependent, they need to coordinate their work often in order avoid
%code duplication and breaking each other's component funcionality.

Therefore, to optimize the development, a development manager can decompose the
system into modules, or groups of components, so that there are as few
inter-module dependencies as possible.% (i.e., the modules are weakly couple.).
Each module is, then, threated as an unit of work assignment \cite{Parnas1972}:
each module is assigned to one team and, because of the little inter-module
dependency, the communication between teams is reduced, therefore speeding up
the development.

Therefore, to decompose a system, a manager needs to have global knowledge of
the dependencies between its components. Although dependency analysis tools can
extract such information from the source code of the system, the task of finding
modules can be overwhelming because of the large number of components and
dependencies.

Software clustering algorithms, also known as architecture recovery algorithms
\cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005},
%\cite{Pollet2007}, 
automatically decompose a software system modules by analyzing the network of
dependencies between its components.  Although many algorithms have been
proposed, there are few comparative studies.
%empirical studies regarding the quality of the decompositions they find.
%EVALUATION CRITERION
One important test for a software clustering algorithm is to apply it to a
collection of software systems with known reference decompositions made by
experienced developers \cite{Anquetil1999}. It is expected that a good algorithm
finds decompositions that are similar to reference decompositions, which can be
measured by a similarity metric such as MoJo \cite{Tzerpos1999} or EdgeSim
\cite{Mitchell2001}.

Unfortunately there are few publicly available reference decompositions
\cite{Koschke2000} and, as a result, few comparative studies exist. Also,
because it is costly to obtain reference decompositions for large systems, most
studies test the algorithms with a couple of small and medium systems
\cite{Anquetil1999,Maqbool2007,Bittencourt2009}.

In our research we propose to evaluate software clustering algorithms by
applying them to synthetic, i.e., computer generated, component dependency
networks with built-in reference decompositions. This approach enables one to
test algorithms with arbitrarily large samples in a controlled manner.

Of course, it is desirable that the synthetic networks resemble networks
extracted from real software systems. In this paper, we present three network
models and show empirically that all three models are capable of synthesizing
networks that resemble software networks. We consider object-oriented systems
with classes as components.

The paper is organized as follows:
Section \ref{sec:models} presents three network models.
%
Section \ref{sec:characterization} demonstrates a method to tell if a network
resembles a software network.
%
Section \ref{sec:evaluation} describes an experiment that shows that the three
models synthesize networks that resemble software networks.
%
Section \ref{sec:conclusion} discusses the impact of this research and future
work.

@@

@@

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, it has been found that many networks
arising from sociology, biology, technology and other domains present remarkable
structural similarities. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to class dependency networks and it was shown
that they are also scale-free networks \cite{Myers2003,Valverde2002}. Thus, the
study of the structure of object-oriented software systems can benefit from the
vast literature available about scale-free networks.

In particular, many scale-free network models were proposed. These models are
algorithms that were proven, either formally or empirically, to synthesize
networks that are scale-free. In this section we present three models that
synthesize scale-free networks with built-in modular decompositions: BCR+, CGW,
and LF. Their output can be used to test software clustering algorithms.

% directed unweighted graphs

%\subsection{ER (Erdos-Renyi)}
%\subsection{Configuration model} %XXX consider using with the most typical sw

@@

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be too low, because most software networks would be misclassified.
So we chose the value 0.88, that is immediatly above the second greater S-score
for a non-software network. With this value, we have both high recall (95.4\%)
and high precision (96.9\%).
